The Weakly Nonlinear Magnetorotational Instability in a local geometry

### 2017-03-08 paper/thingap_ref_response/response_to_referee.tex

\item A more extensive discussion of the role of the neutral mode with amplitude $\beta$ that both URM and the present paper argue should be kept, but, at least to the order in $\epsilon$ considered, satisfies an amplitude equation that is completely decoupled from the MRI mode, with amplitude $\alpha$. See the comment on coupling of damped modes above, which may be related to this issue.

\textbf{The neutral mode is a spatially constant azimuthal magnetic field. It cannot nonlinearly interact with any modes -- damped or otherwise -- because all of the nonlinearities involve derivatives. This remains true when ambipolar diffusion is considered. We have added two sentences to Section 3 to clarify this point.}

%Neutral mode doesn’t exist in cylindrical geometry, yet we find same GLE -- can we just argue that it’s not important for saturation? 
%The beta mode arises because a constant By solves the equations. However, this cannot non-linearly couple to any other modes, because all NL terms in the problem involve derivatives. Thus, there is no coupling to it, even though we discard couplings between the other damped eigenmode families in the thin-gap MHD system (that is, the non-MRI modes)

\item The saturated Bz profile shown in Figure 5 of Paper I is intriguing in light of the fact that the authors achieved instability by dialing down Bz from the marginally stable value. Would a radial average of the new Bz, which is stronger at the edges and weaker in the middle, reveal that it is marginally stable? Do the results bear on whether the MRI is a dynamo? There is no counterpart figure in Paper II; the authors should comment on dynamo action here, too.

\textbf{Yes, the saturated Bz is marginally stable, and we have added a sentence to that effect in the text. We cannot comment on dynamo action because our setup is axisymmetric. }

\end{itemize}

\item A small point – the parameter called $\beta$ is not the standard definition; why not introduce the Cowling number?

\item Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of $\epsilon$? Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.					\end{enumerate}

### 2017-03-08 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
more response, Bz and no dynamo comment

Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance ($\partial_x$ of Equation~\ref{eqset4} with $\partial_t B_{y} = 0$). Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers. Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition. \citet{Ebrahimi:2009ey} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean. They considered two important cases, one in which the mean flow was forced to remain at its initial, quasi-Keplerian state for all time, and one in which the background flow was allowed to evolve. This is a crucial difference between the shearing box and our narrow-gap TC flow: perturbations in our simulation can adjust the background flow, whereas in a shearing box, the shear periodicity forbids perturbations from affecting the mean flow. In the case with a freely evolving background flow, \citet{Ebrahimi:2009ey} found a saturated state quite similar to ours: field pushed to the boundaries, and a reduction in shear in the bulk of the flow. Their flows have less pronounced boundary layers, likely because of their much larger $\Pm = 0.1 -1$.

@@

@@

\bibitem[{Kunz \& Balbus(2004)}]{Kunz:2004ib}
Kunz, M. W. \& Balbus, S. A. 2004, MNRAS 348, 355

\bibitem[{Kunz \& Lesur(2013)}]{Kunz:2013}
Kunz, M. W. \& Lesur, G. 2013, MNRAS 434, 2295

\bibitem[{Latter et al.(2015)}]{Latter:2015}
Latter, H.N., Fromang, S., \& Faure, J., 2015, MNRAS 453, 3257

\bibitem[{Latter et al.(2010)}]{Latter:2010iz}
Latter, H.N., Fromang, S., \& Gressel, O., 2010, MNRAS 406, 848

\bibitem[{Makwana et al.(2011)}]{Makwana:2011}
Makwana, K.D., Terry, P.W., Kim, J.H., \& Hatch, D.R. 2011, \pop, 18, 1

>>>>

Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance ($\partial_x$ of Equation~\ref{eqset4} with $\partial_t B_{y} = 0$). Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers. %Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition.
\edit1{The radial average of the saturated $B_z$ is $B_0$, i.e. $B_z$ is marginally stable.}
 \citet{Ebrahimi:2009ey} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean. They considered two important cases, one in which the mean flow was forced to remain at its initial, quasi-Keplerian state for all time, and one in which the background flow was allowed to evolve. This is a crucial difference between the shearing box and our narrow-gap TC flow: perturbations in our simulation can adjust the background flow, whereas in a shearing box, the shear periodicity forbids perturbations from affecting the mean flow. In the case with a freely evolving background flow, \citet{Ebrahimi:2009ey} found a saturated state quite similar to ours: field pushed to the boundaries, and a reduction in shear in the bulk of the flow. Their flows have less pronounced boundary layers, likely because of their much larger $\Pm = 0.1 -1$.

@@

\clearpage
\appendix

\section{A. Detailed Equations}\label{app:matrices}

Here we detail the perturbation analysis described in Section \ref{sec:wnl}. The perturbation series is described by Equations \ref{eq:ordere} - \ref{eq:ordere3}, where

\beq
\mathcal{L} = \mathcal{L}_0 + \mathcal{L}_1 \partial_z + \mathcal{L}_2 \partial_z^2 + \mathcal{L}_3 \partial_z^3 + \mathcal{L}_4 \partial_z^4,
\eeq

\beq
\widetilde{\mathcal{L}}_1 =  \mathcal{L}_1 + 2\mathcal{L}_2\partial_z + 3\mathcal{L}_3\partial_z^2 + 4\mathcal{L}_4\partial_z^3
\eeq

\beq
\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2
\eeq

\beq
\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
\eeq

and the constituent matrices are defined as

\beq
\mathcal{D} = \left[\begin{matrix}
\nabla^2 & 0 & 0 & 0 \\
0 & 1& 0 & 0 \\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1 \\
\end{matrix}\right]
\eeq

\beq
\mathcal{L}_0 = \left[\begin{matrix}
-\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
0 & -\frac{1}{\reye}\partial_x^2 & 0 &0 \\
0 & 0 & -\frac{1}{\reym}\partial_x^2 & 0 \\
0 & 0 & 0 & -\frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right]
\eeq

\beq
\mathcal{L}_1 = \left[\begin{matrix}
0 & -2 & -\frac{2}{\beta}\partial_x^2 & 0 \\
(2-q)\Omega_0 & 0 & 0 & -\frac{2}{\beta} \\
-1 & 0 & 0 & 0 \\
0 & -1 & q\Omega_0 & 0 \\ \end{matrix}\right] 
\eeq

\beq
\mathcal{L}_2 = \left[\begin{matrix}
-2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
0 & -\frac{1}{\reye} & 0 & 0 \\
0 & 0 & -\frac{1}{\reym} & 0 \\
0 & 0 & 0 & -\frac{1}{\reym} \\ \end{matrix}\right]
\eeq

\beq
\mathcal{L}_3 = \left[\begin{matrix}
0 & 0 & -\frac{2}{\beta} & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\ \end{matrix} \right]
\eeq

\beq
\mathcal{L}_4 = \left[\begin{matrix}
-\frac{1}{\reye} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\ \end{matrix}\right] 
\eeq

\beq
\mathcal{G} = \left[\begin{matrix}
0 & 0 & -\frac{2}{\beta}\partial_x^2 & 0 \\
0 & 0 & 0 & -\frac{2}{\beta} \\
-1 & 0 & 0 & 0 \\
0 & -1 & 0 & 0 \\ \end{matrix} \right]
\eeq

Once perturbed, the system is solved for successive orders of $\epsilon$ (Equations \ref{eq:ordere} - \ref{eq:ordere3}). $\mathcal{O}(\epsilon)$ is the linear system. At $\mathcal{O}(\epsilon^2)$, first-order MRI modes nonlinearly interact with themselves and with their complex conjugates, and so the term $\mathbf{N}_2$ in Equation \ref{eq:ordere2} has the form

\beq
\mathbf{N}_2 = |\alpha|^2 \mathbf{N}_{20} + \alpha^2 \mathbf{N}_{22} e^{2 i k_c z}
\eeq

(see Appendix \ref{app:nonlinear_terms} for the full form of $\mathbf{N}_{20}$ and $\mathbf{N}_{22}$).

Note that, following the notation of \citet{Umurhan:2007hs}, the subscripts refer to $\epsilon$ order, $z$ order, successively, such that $\mathbf{N}_{22}$ is the second-order nonlinear term which corresponds to $e^{2 i k_c z}$ z-dependence.

Equation \ref{eq:ordere2} is solved as three separate systems of equations, one for each possible $z$ resonance:

\begin{align}
\mathcal{L}\mathbf{V}_{20} & = \mathbf{N}_{20}\\
\mathcal{L}\mathbf{V}_{21} & = - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_{11} \\
\mathcal{L}\mathbf{V}_{22} & = \mathbf{N}_{22}
\end{align}

Finally, at $\mathcal{O}(\epsilon^3)$ we eliminate secular terms to close the system. Secular terms are terms which are resonant with the solution to the homogenous linear equation (Equation \ref{eq:ordere}), and which cause the higher-order solutions to grow without bound. The solvability criterion we enforce to eliminate these terms is the vanishing of the inner product of the solution to the adjoint linear homogenous equation $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ with the nonhomogenous terms in Equation \ref{eq:ordere3}, namely

\beq\label{eq:innerprod_gle}
\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2.
\eeq

This solvability criterion derives from a corollary to the Fredholm Alternative (see Paper II for a formal definition).

Equation \ref{eq:innerprod_gle} can be rewritten as Equation \ref{eq:gle}, the Ginzburg-Landau equation, where the coefficients are

\beq
b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
\eeq

\beq
h = \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
\eeq

and

\beq
c = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle. 
\eeq

We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as

\beq
\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \langle \mathcal{L}^\dagger \mathbf{V}^\dagger | \mathbf{V} \rangle,
\eeq

where the inner product is defined as

\beq\label{eq:inner_product_def}
\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{x_1}^{x_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} x \mathrm{d} z.
\eeq

The solution to the adjoint homogenous equation has the form

\beq
\mathbf{V^\dagger} = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.
\eeq

As noted by URM07, a second amplitude equation for a spatially constant azimuthal magnetic field mode arises from the terms in the $\mathcal{O}(\epsilon^3)$ equation which contain no $z$ dependence. This is a diffusion equation, so the neutral mode simply decays away.

\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}

At each order in our perturbation series, lower-order MRI modes nonlinearly interact. Thus there is a nonlinear term contribution at $\mathcal{O}(\epsilon^2)$ and $\mathcal{O}(\epsilon^3)$. Here we detail the form of these nonlinear terms.

The overall nonlinear contribution to our system, written as a vector $\mathbf{N}$ in Equation \ref{eq:unperturbed_matrix_equations}, is

\beq
\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
\eeq

where

\beq
N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
\eeq

\beq
N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
\eeq

\beq
N_2^{(A)} = - J(A_1, \Psi_1) 
\eeq

\beq
N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)
\eeq

and

\beq
\begin{split}
N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
\end{split}
\eeq

\beq
N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
\eeq

\beq
N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1)
\eeq

\beq
N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
\eeq

$\mathbf{N_2}$ and $\mathbf{N_3}$ expand to become

\beq
\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
\eeq

and

\beq
\mathbf{N_3} = \alpha^3 \mathbb{N}_{33} e^{i 3 k_c z} + \alpha\partial_Z\alpha \mathbb{N}_{32} e^{i 2 k_c z} + \alpha\left|\alpha\right|^2 \mathbb{N}_{31} e^{i k_c z} + \alpha \partial_Z  \beta \mathbb{\widetilde{N}}_{31} e^{i k_c z} + \alpha^*\partial_Z \alpha \mathbb{N}_{30} + c.c.
\eeq

The second order nonlinear terms are \\

\beq
\begin{split}
N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
\end{split}
\eeq

\beq
\begin{split}
N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
\end{split}
\eeq

\beq
N_{22}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i k_c \Psi_{11}
\eeq

\beq
N_{22}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i k_c B_{11} - i k_c A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i k_c u_{11}
\eeq

\beq
\begin{split}
N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
\end{split}
\eeq

\beq
N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
\eeq

\beq
N_{20}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i k_c \Psi_{11}^*
\eeq

\beq
N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
\eeq \\

and the third order nonlinear terms become \\

\beq
\begin{split}
N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
& + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
& - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
\end{split}
\eeq

\beq
\begin{split}
N_{31}^{(u)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
& - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
\end{split}
\eeq

\beq
\begin{split}
N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
 + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
 & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
\end{split}
\eeq

\beq
\begin{split}
N_{31}^{(B)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 k_c\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
& - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
& - i k_c \left(A_{11} \cdot \partial_x u_{20}\right) - i k_c \left(A_{11} \cdot \partial_x u_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
& i k_c \left(u_{11} \cdot \partial_x A_{20} \right) + i k_c \left(u_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 k_c \left(\partial_x u_{11}^* \cdot A_{22}\right)
\end{split}
\eeq

\section{C. Linear dispersion relation}\label{app:dispersion}

The linear dispersion relation, which determines the variable scalings in the multiple scales analysis. This relation is found by perturbing the linear system (Equation \ref{eq:ordere}) with a small perturbation of the form $e^{\sigma t + i k_x x + i k_z z}$. Note that the spatial eigenvalues appear as $k_z^2$ and $k_x^2$ at lowest order.

\beq
\begin{split}
& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
\end{split}
\eeq

@@

\bibitem[{Kunz \& Balbus(2004)}]{Kunz:2004ib}
Kunz, M. W. \& Balbus, S. A. 2004, \mnras 348, 355

\bibitem[{Kunz \& Lesur(2013)}]{Kunz:2013}
Kunz, M. W. \& Lesur, G. 2013, \mnras 434, 2295

\bibitem[{Latter et al.(2015)}]{Latter:2015}
Latter, H.N., Fromang, S., \& Faure, J., 2015, \mnras 453, 3257

\bibitem[{Latter et al.(2010)}]{Latter:2010iz}
Latter, H.N., Fromang, S., \& Gressel, O., 2010, \mnras 406, 848

\bibitem[{Makwana et al.(2011)}]{Makwana:2011}
Makwana, K.D., Terry, P.W., Kim, J.H., \& Hatch, D.R. 2011, Phys. of Plasmas, 18, 1
